\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-11-19 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Encore des changements
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>

Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019. Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs. While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis (\cites{samuelson1954}),\footnote{``Iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.} a recent strand of the empirical literature points to the existence of an additive component, dependent on the physical quantities traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}). Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip

@@

To what extent should we care about additive costs? Our paper answers this question from two different perspectives.
On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019. By allowing for country-product-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel. Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows. Importantly, the difference can be attributed to our more flexible modeling of the additive component. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to around 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.13 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
Expressed in terms of the export price, the additive costs amount to 2.8\% and 1.7\% for ocean shipping and air transport respectively.
This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3\% and 2.4\% for vessel and air shipping respectively (on average over the period).
Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as displayed in the last rows of Table \ref{tab:summary_results}. The per-kg cost dimension is quantitatively sizable.

@@

Our estimation results on the size of additive transport costs can be compared to \cite{Irrazabal_2015}, the most recent evidence on this question.
Based on Norwegian trade data for 2004, \cite{Irrazabal_2015} estimate that additive costs amount to 6\% of export price multiplied by the ad-valorem cost, i.e.
$\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
This allows building the ratio in similar terms to them, allowing us to gain in generality in this respect.
Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping and 1.5\% in air transport.
This may seem surprisingly low, in contrast to the 6\% obtained by \cite{Irrazabal_2015}.
This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
While the database of \cite{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
\cites{Irrazabal_2015} results on the year 2004 imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
For the same export price multiplied by the ad-valorem cost of 100 USD, our results for 2004 yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.

@@

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in vessel shipping. Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
Further, the transport cost reduction is much smoother in vessel transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing). Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
This is particularly the case for the multiplicative component (panel (d)).\smallskip

@@

The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping. If anything, we find the opposite result here: trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern. This drives us to investigate this difference further.

@@

In line with the literature, we find that a reduction of trade costs  induces welfare gains.
% whatever the transport mode (and the associated magnitude of the transport costs decrease).
The originality of our results rather rests on comparing the welfare gains depending on the type of cost reduction.
Welfare gains are higher when the trade cost reduction includes a decrease in additive trade costs.
This is consistent with the analytical results of \cite{sorensen2014} and the quantitative ones of \cite{Irrazabal_2015}.
%in a different exercise of trade liberalization though. \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness. Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization. Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.}
Quantitatively, the welfare gains are sizeable. In our preferred scenario, the reduction in variable transport costs in Air transport generates a 40\% higher welfare gain  when such a reduction partly occurs through the additive component, rather than when it happens through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)). In maritime transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%. This can be set in connection with the fact that in maritime transport, ad-valorem costs have reduced more than the additive ones, implying a larger $\beta$ at the end of the period.\footnote{See Appendix \ref{app:theoretical_model} for a similar result.}
%Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem. If we share a similar conclusion regarding the importance of additive costs, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
\smallskip

@@

The welfare-dampening impact of the variety effect is also stronger under this scenario. Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the
share of exporters. As the selection on the domestic market is also more stringent, the
mass of total active firms reduces by a larger amount. From the consumer side, this induces
a more negative variety effect that is welfare-detrimental everything else equal. As for the
pure-ad-valorem case though, the strength of the (larger) competitiveness effect dominates,
inducing positive welfare gains, and even larger when the trade liberalization involves a
reduction in additive costs.\medskip

\paragraph{Taking into account the reduction in fixed costs} As mentioned above, our preferred scenario includes, in addition to the reduction of variable costs, a reduction in fixed export costs, based on \cite{Lincoln_McCallum2018}'s empirical evidence.
The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (2).
Two results emerge.
First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
While this is not a surprising result \textit{per se} (we now consider two sources of international frictions, rather than one), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, higher by around 50\% relative to the case when only the two components of the variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively. Put differently, the reduction of the sunk export cost accounts for the majority of welfare gains. This is not the case under additive variable costs; in this case, the reduction of the distortion due to the additive costs continues to substantially matter in the welfare-enhancing effect of the globalization process.}  This can be accounted for as follows. The reduction in the fixed export cost pushes more firms to export everything else equal. This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market. As a result, the joined reduction in variable costs and in $f_x$ dampens the negativity of the variety effect (which even turns out to be positive in some cases). This, combined with a positive price competitiveness effect, accounts for the stronger welfare gains reported in Panel (2).

@@

In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results emphasize the importance of taking into account the additive components of international trade costs. Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods. This, in turn, has important consequences in normative terms. Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 60\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.66\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.6\%). A similar result applies with the observed 57\% reduction in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world). Neglecting the existence and long-term evolution additive costs thus leads to a substantial underestimation of the welfare gains of trade cost reduction.

@@

This paper assesses the magnitude and the impact of additive international transport costs. Empirically, we exploit the differences between the import and the export prices taken from the US import database over 1974-2019, to estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean). We find that additive costs are quantitatively sizable: On average over the period, they amount to 2.8\% of the export price unit values for ocean shipping, and 1.7\% for air transport, accounting for 30\% and 45\% of total transport costs in air and maritime transport respectively.

To what extent should we care about additive costs? Our paper addresses this question through two different angles. First, we show the importance of integrating the additive component in accounting for the time trend of international transport costs observed over 1974-2019. Allowing for a varying share of additive costs in the product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role. If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in maritime transport.

@@

\section{Data Appendix \label{app:data}}

@@

In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2019, when the degree of classification retained ($s$) is at the 3-digit classification level (products are at 5-digit level).
Table \ref{tab:result_air_3d_detail} reports the results for air transport.
The results for vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.

@@

Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
On average over the whole period, the $R^{2}$ doubles when per-kg costs are included for air, and increases by 50\% for vessel.
Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes.
Both AIC and log-likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
These results holds whatever the considered year. \footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix B.2. If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification.}
%This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
%When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
%In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
%Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.

@@

Table \ref{tab: FS_sitc5} suggests that custom duties variations are a satisfactory predictor of fas prices $\widetilde{p}_{ik}$, with strong significance and the correct, negative sign. The lagged fas price $\log \widetilde{p}_{ikt-1}$ displays also the expected positive sign: with a $\beta$ parameter around 0.75 on average, fas prices exhibit substantial persistence. The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation. In addition, Figure \ref{fig: FS_IV_SITC5} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a very small number of occasions, corresponding to abrupt fluctuations in oil prices. This is typically the case in 1975, with $\alpha$=-1.83. All in all, these estimates make us confident in the validity of our instrumentation strategy.

@@

Results are qualitatively very similar to those reported on the longer-time-span, more-aggregated-SITC5 dataset previously reported. The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average close to 12, slightly higher than in our benchmark. As previously, Figure \ref{fig: FS_IV_HS10} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a few years which seem again to coincide with significant fluctuations in oil prices. Interestingly, the value of the $\alpha$ parameter is on average -0.36, lower than in our benchmark on SITC5 data. At this highly disaggregated level, pricing-to-market behavior seems weaker, indicating that the per-product average firm is smaller than the average per-sector one. In addition, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006. In any case, this strengthens our confidence in the validity of our estimation strategy.

>>>>

Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019. Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs. While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis (\cites{samuelson1954}),\footnote{``Iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.} a recent strand of the empirical literature points to the existence of an additive component, dependent on the physical quantities traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}). Based on US sectoral data, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or vessel transport: only waterborne and airborne shipments are included).\smallskip

@@

To what extent should we care about additive costs? Our paper answers this question from two different perspectives.
On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019. By allowing for country-product-time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel. Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in vessel transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows. Importantly, the difference can be attributed to our more flexible modeling of the additive component. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs.
On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to around 10 cents (0.1 USD) per kilogram exported in vessel transport, and 1.13 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case.
Expressed in terms of the export price, the additive costs amount to 2.8\% and 1.7\% for ocean shipping and air transport respectively.
This is slightly lower than the estimates for the ad-valorem component, which is estimated at 3\% and 2.4\% for vessel and air shipping respectively (on average over the period).
Put differently, the additive component amounts to 44\% of total transport costs in vessel, and 30\% in Air, as displayed in the last rows of Table \ref{tab:summary_results}. The per-kg cost dimension is quantitatively sizable.

@@

Our estimation results on the size of additive transport costs can be compared to \cite{Irrazabal_2015}, the most recent evidence on this question.
Based on Norwegian trade data for 2004, \cite{Irrazabal_2015} estimate that additive costs amount to 6\% of export price multiplied by the ad-valorem cost, i.e.
$\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results).
One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem).
This allows building the ratio in similar terms to them, allowing us to gain in generality in this respect.
Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping and 1.5\% in air transport.
This may seem surprisingly low, in contrast to the 6\% obtained by \cite{Irrazabal_2015}.
This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case.
While the database of \cite{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases).
In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment.
\cites{Irrazabal_2015} results on the year 2004 imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs.
For the same export price multiplied by the ad-valorem cost of 100 USD, our results for 2004 yield the value of 2.8 (1.5) USD paid in additive transport costs for vessel (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuild the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$.
Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 25\% (air) and 47\% (vessel) of additive trade costs are attributable to international transport.

@@

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in vessel shipping. Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)).
Further, the transport cost reduction is much smoother in vessel transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing). Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
Considering the raw series (plain line), vessel transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
This is particularly the case for the multiplicative component (panel (d)).\smallskip

@@

The last results stand in sharp contrast with \cite{hummels2007}, who finds that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and vessel shipping. If anything, we find the opposite result here: trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward pattern. This drives us to investigate this difference further.

@@

In line with the literature, we find that a reduction of trade costs  induces welfare gains.
% whatever the transport mode (and the associated magnitude of the transport costs decrease).
The originality of our results rather rests on comparing the welfare gains depending on the type of cost reduction.
Welfare gains are higher when the trade cost reduction includes a decrease in additive trade costs.
This is consistent with the analytical results of \cite{sorensen2014} and the quantitative ones of \cite{Irrazabal_2015}.
%in a different exercise of trade liberalization though. \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness. Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization. Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.}
Quantitatively, the welfare gains are sizeable. In our preferred scenario, the reduction in variable transport costs in Air transport generates a 40\% higher welfare gain  when such a reduction partly occurs through the additive component, rather than when it happens through a decrease in ad-valorem costs only (Table \ref{tab:resultats_modele}, panel (2), Column (B) vs (A)). In vessel transport, the extra welfare gains due to the joined reduction of both transport costs components are more modest, with a magnitude of order of 14\%. This can be set in connection with the fact that in vessel transport, ad-valorem costs have reduced more than the additive ones, implying a larger $\beta$ at the end of the period.\footnote{See Appendix \ref{app:theoretical_model} for a similar result.}
%Closely related to our work, \cite{Irrazabal_2015} also quantify the welfare losses of imposing trade taxes depending on whether they are additive or ad-valorem. If we share a similar conclusion regarding the importance of additive costs, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem).
\smallskip

@@

The welfare-dampening impact of the variety effect is also stronger under this scenario. Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the
share of exporters. As the selection on the domestic market is also more stringent, the
mass of total active firms reduces by a larger amount. From the consumer side, this induces
a more negative variety effect that is welfare-detrimental everything else equal. As for the
pure-ad-valorem case though, the strength of the (larger) competitiveness effect dominates,
inducing positive welfare gains. These gains are larger when the trade cost reduction involves a
reduction in additive costs.\medskip

\paragraph{Taking into account the reduction in fixed costs} As mentioned above, our preferred scenario includes, in addition to the reduction of variable costs, a reduction in fixed export costs, based on \cite{Lincoln_McCallum2018}'s empirical evidence.
The associated welfare gains are reported in Table \ref{tab:resultats_modele}, Panel (2).
Two results emerge.
First, the welfare gains induced by the trade cost reduction are stronger in magnitude.
While this is not a surprising result \textit{per se} (we now reduce two sources of international frictions, rather than one), this exercise delivers a quantification of the welfare gains associated with a reduction in the fixed export cost. They are substantial, higher by around 50\% relative to the case when only variable transport costs decrease.\footnote{Under the scenario with only ad-valorem costs (Columns (A) and (C)), the extra welfare gains due to the reduction of $f_x$ are much higher, by 66\% (Vessel) and 89\% (Air) respectively. Put differently, the reduction of the sunk export cost accounts for the majority of welfare gains. This is not the case under additive variable costs; in this case, the reduction of the distortion due to the additive costs continues to substantially matter in the welfare-enhancing effect of the globalization process.}  This can be accounted for as follows. The reduction in the fixed export cost pushes more firms to export, everything else being equal. This increases the number of available imported varieties for the consumer, which counteracts the reduction in the number of domestic varieties induced by the tougher selection on the domestic market. As a result, the joined reduction in variable costs and in $f_x$ dampens the negativity of the variety effect (which even turns out to be positive in some cases). This, combined with a positive price competitiveness effect, accounts for the stronger welfare gains reported in Panel (2).

@@

In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results emphasize the importance of taking into account the additive components of international trade costs. Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods. This, in turn, has important consequences in welfare terms. Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 60\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.66\% increase in welfare, much higher than if only ad-valorem costs were modeled (2.6\%). A similar result applies with the observed 57\% reduction in total variable costs in vessel transport along with the reduction in the fixed export costs, with a relative increase in welfare by 3.2\% (as compared to 2.8\% in a pure ad-valorem world). Neglecting the existence and long-term evolution additive costs thus leads to a substantial underestimation of the welfare gains of trade cost reduction.

@@

This paper assesses the magnitude and the impact of additive international transport costs. Empirically, we exploit the differences between the import and the export prices taken from the US import database over 1974-2019, to estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean). We find that additive costs are quantitatively sizable: On average over the period, they amount to 2.8\% of the export price unit values for ocean shipping, and 1.7\% for air transport, accounting for 30\% and 45\% of total transport costs in air and vessel transport respectively.

To what extent should we care about additive costs? Our paper addresses this question through two different angles. First, we show the importance of integrating the additive component in accounting for the time trend of international transport costs observed over 1974-2019. Allowing for a varying share of additive costs in the product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role. If anything, trade composition effects have contributed to amplifying the reduction in the pure transport costs in vessel transport.

@@

\section{Data Appendix: customs value and import charges\label{app:data}}

@@

In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit).
Specifically, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2019, when the degree of classification retained ($s$) is at the 3-digit classification level (products are at 5-digit level).
Table \ref{tab:result_air_3d_detail} reports the results for air transport.
The results for vessel transport are displayed in Table \ref{tab:result_ves_3d_detail}.
Full annual data are available in the Online Appendix.

@@

Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel} lead to the same conclusion: The inclusion of the additive term leads to an improvement of the quality of fit, whatever the considered criterion or the transport mode.
On average over the whole period, the $R^{2}$ doubles when additive costs are included for air, and increases by 50\% for vessel.
Similar qualitative conclusions arise from the comparisons of the standard errors of the regression (SER).
Regarding the other criteria, improvements allowed by the inclusion of the additive term are roughly of the same extent across transport modes.
Both AIC and log-likelihood statistics decrease with the inclusion of the additive term, and the log-likelihood test unambiguously rejects the null of statistical equivalence of the two models.
These results holds whatever the considered year. \footnote{For comparison purposes, we provide a similar goodness-of-fit exercise at the 4-digit product level (4-digits), reported in Online Appendix B.2. It is slightly more favorable to the model including additive costs.}
%This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
%When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
%In other words, if using a more disaggregated classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.
%Further, the same conclusion established at the 3-digit level regarding the significant role of the additive component in fitting international transport costs emerges at the 4-digit level.

@@

Table \ref{tab: FS_sitc5} suggests that custom duties variations are a satisfactory predictor of fas prices $\widetilde{p}_{ik}$, with strong significance and the correct, negative sign. The lagged fas price $\log \widetilde{p}_{ikt-1}$ displays also the expected positive sign: with a $\beta$ parameter around 0.75 on average, fas prices exhibit substantial persistence. The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average above 10, the ``rule-of-thumb'' threshold value recommended by \cite{staiger_stock} as a test for weak instrumentation. It is true that the median is at 6.63, indicating that estimates may be noisier for more than half of the considered years. That said, the use of the lagged fas price as additional instrument ensures an overall strong predictive power of the first-stage equation, and consequently, accurate (i.e. less noisy) second-stage estimates.

In addition, Figure \ref{fig: FS_IV_SITC5} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a very small number of occasions, corresponding to abrupt fluctuations in oil prices. This is typically the case in 1975, with $\alpha$=-1.83. All in all, these estimates make us confident in the validity of our instrumentation strategy.

@@

Results are qualitatively very similar to those reported on the longer-time-span, more-aggregated-SITC5 dataset previously reported. The F-statistic on the excluded instrument ($\frac{\Delta \tau^d_{is(k)t}}{1+\tau_{is(k)t-1}^d}$ ($\widehat{\alpha}$)) is on average close to 12, with a median closer to 10. In both cases, these numbers are slightly higher than in our benchmark.

As previously, Figure \ref{fig: FS_IV_HS10} shows that most values of the $\alpha$ parameter stand between 0 and -1, as expected, except in a few years which seem again to coincide with significant fluctuations in oil prices. Interestingly, the value of the $\alpha$ parameter is on average -0.36, lower than in our benchmark on SITC5 data. At this highly disaggregated level, pricing-to-market behavior seems weaker, indicating that the per-product average firm is smaller than the average per-sector one. In addition, this average value is almost identical to the one found by \citet{Fontagne_Martin_Orefice2018} on French HS6 data between 1995 and 2010 (see the first-stage coefficients in their table 6), or by \citet{Bussiere2013} on country-level data between 1980 and 2006. In any case, this strengthens our confidence in the validity of our estimation strategy.
